```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(tidymodels)
library(ranger)


```


read in data


```{r}
dat <- combined.mir1

```
check str
```{r}
str(dat)

table(dat$STATUS)

```

## split data
```{r}
set.seed(525)
omic_split <- dat %>%
  initial_split(prop = 0.8, strata = STATUS)
omic_train <- training(omic_split)
omic_test <- testing(omic_split)
#ready <- omic_test
#save(ready, file="ready.rda")
```


## preprocessing
```{r}
omic_recipe <- recipe(STATUS ~ ., data = omic_train) %>% 
  step_normalize(all_numeric_predictors()) %>% 
 # step_dummy(all_nominal_predictors(),- all_outcomes()) %>% 
  step_upsample(STATUS, over_ratio = 1) %>% 
  step_nzv(all_predictors()) %>% 
  step_zv(all_predictors()) %>% 
 # step_corr(all_predictors()) %>%  
  prep()
omic_recipe
```


```{r}
omic_train_baked <- bake(omic_recipe, NULL)
omic_test_baked <- bake(omic_recipe, new_data = omic_test)
table(omic_train_baked$STATUS)
table(omic_test_baked$STATUS)
```


## random forest

### model specification and workflow creation
```{r}
rf <- rand_forest(mtry = tune(), 
                  trees = tune(), 
                  min_n = tune()) %>% 
  set_mode("classification") %>% 
  set_engine("ranger")
rf_wf <- 
  workflow() %>% 
  add_model(rf) %>% 
  add_recipe(omic_recipe) 
```


### prepare for hyperparameter tuning
```{r}
rf_grid <- grid_regular(mtry(range = c(5L,15L)),
                        min_n(range = c(5, 10)),
                        trees(range = c(1000L, 2000L)),
                             levels = 2)
set.seed(525)
omic_folds <- vfold_cv(omic_train, v = 5, repeats = 1)
```


### train and tune model (with tune_grid)
```{r}
doParallel::registerDoParallel(cores = 5)
metrics = metric_set(roc_auc, accuracy)
rf_res <- 
  rf_wf %>% 
  tune_grid(
    resamples = omic_folds,
    grid = rf_grid, 
    metrics = metrics,
    control = control_grid(save_pred = TRUE)
    )
doParallel::stopImplicitCluster()
foreach::registerDoSEQ()
```


### evaluate the models
```{r}
rf_res %>% 
  collect_metrics() %>% 
  ggplot(aes(x = mtry, y = mean, color = .metric)) + 
  geom_point() + 
  geom_line() + ylab("") +
  facet_wrap(~trees + min_n) +
  ggtitle(label = "model performance", 
          subtitle = "random forest") 
rf_best <- rf_res %>%
  select_best(metric = "roc_auc")
rf_auc <- 
  rf_res %>% 
  collect_predictions(parameters = rf_best) %>% 
  roc_curve(truth = STATUS, estimate = .pred_no) %>% 
  mutate(model = "random forest")
autoplot(rf_auc)
```

### train and fit the best model
```{r}

set.seed(567)
final_rf_wf <- 
  rf_wf %>% 
  finalize_workflow(parameters = rf_best)
#saveRDS(final_rf_wf,file="final_rf_wf.RDS")
final_rf_fit <- 
  final_rf_wf %>%
  last_fit(omic_split, 
           metrics = metrics) 
final_rf_fit %>%
  collect_metrics()
final_rf_fit %>%
  collect_predictions() %>% 
  roc_curve(truth = STATUS, estimate = .pred_no) %>% 
  autoplot()
```

https://www.rebeccabarter.com/blog/2020-03-25_machine_learning/#variable-importance
```{r}

test_predictions <- final_rf_fit %>% collect_predictions()
test_predictions

```

```{r}
# generate a confusion matrix
test_predictions %>% 
  conf_mat(truth = STATUS, estimate = .pred_class)

```
```{r}

test_predictions %>%
  ggplot() +
  geom_density(aes(x = .pred_no, fill = STATUS), 
               alpha = 0.5)

```
```{r}
final_model <- fit(final_rf_wf, dat)
final_model
```

```{r}
ranger_obj <- pull_workflow_fit(final_model)$fit
ranger_obj

```
```{r}
ranger_obj$variable.importance
```

## logistic regression

### model specification and workflow creation
```{r}
log_reg <- logistic_reg(penalty = tune(), mixture = tune()) %>% 
  set_mode(mode = "classification") %>% 
  set_engine(engine = "glmnet")
lr_wf <- workflow() %>% 
  add_recipe(omic_recipe) %>% 
  add_model(log_reg)
```


### prepare for hyperparameter tuning
```{r}
lr_param <- parameters(log_reg)
lr_grid <- grid_regular(penalty(), mixture(), levels = 3)
set.seed(525)
omic_folds <- vfold_cv(omic_train, v = 5, repeats = 1)
```


### train and tune model (with tune_grid)
```{r}
parallel::detectCores(logical = FALSE)
doParallel::registerDoParallel(cores = 5)
metrics = metric_set(roc_auc, sens, spec,accuracy)
lr_res <- 
  lr_wf %>% 
  tune_grid(
    resamples = omic_folds,
    grid = lr_grid, 
    metrics = metrics,
    control = control_grid(save_pred = TRUE)
    )
doParallel::stopImplicitCluster()
foreach::registerDoSEQ()
```

### evaluate the models
```{r}
lr_res %>% 
  collect_metrics() %>% 
  ggplot(aes(x = penalty, y = mean, color = .metric)) + 
  geom_point() + 
  geom_line() + ylab("") +
  facet_wrap(~mixture) +
  scale_x_log10(labels = scales::label_scientific()) +
  ggtitle(label = "model performance", 
          subtitle = "logistic regression") 
lr_best <- lr_res %>%
  select_best(metric = "roc_auc")
lr_auc <- 
  lr_res %>% 
  collect_predictions(parameters = lr_best) %>% 
  roc_curve(truth = STATUS, estimate = .pred_no) %>% 
  mutate(model = "Logistic Regression")
autoplot(lr_auc)
```


### train and fit the best model
```{r}
final_lr_wf <- 
  lr_wf %>% 
  finalize_workflow(parameters = lr_best)
final_lr_fit <- 
  final_lr_wf %>%
  last_fit(omic_split, 
           metrics = metrics) 
final_lr_fit %>%
  collect_metrics()
final_lr_fit %>%
  collect_predictions() %>% 
  roc_curve(truth = STATUS, estimate = .pred_no) %>% 
  autoplot()
#saveRDS(final_lr_wf,file="final_lr_wf.RDS")
```


## knn

### model specification and workflow creation
```{r}
knn <- nearest_neighbor(neighbors = tune()) %>% 
  set_mode("classification") %>% 
  set_engine("kknn")
k_wf <- 
  workflow() %>% 
  add_model(knn) %>% 
  add_recipe(omic_recipe) 
```


### prepare for hyperparameter tuning
```{r}
k_param <- parameters(knn)
k_grid <- grid_regular(neighbors(), levels = 3)
set.seed(525)
omic_folds <- vfold_cv(omic_train, v = 5, repeats = 1)
```


### train and tune model (with tune_grid)
```{r}
doParallel::registerDoParallel(cores = 5)
metrics = metric_set(roc_auc, accuracy)
k_res <- 
  k_wf %>% 
  tune_grid(
    resamples = omic_folds,
    grid = k_grid, 
    metrics = metrics,
    control = control_grid(save_pred = TRUE)
    )
doParallel::stopImplicitCluster()
foreach::registerDoSEQ()
```


### evaluate the models
```{r}
k_res %>% 
  collect_metrics() %>% 
  ggplot(aes(x = neighbors, y = mean, color = .metric)) + 
  geom_point() + 
  geom_line() + ylab("") +
  ggtitle(label = "model performance", 
          subtitle = "knn") 
k_best <- k_res %>%
  select_best(metric = "roc_auc")
k_auc <- 
  k_res %>% 
  collect_predictions(parameters = k_best) %>% 
  roc_curve(truth = STATUS, estimate = .pred_no) %>% 
  mutate(model = "knn")
autoplot(k_auc)
```


### train and fit the best model

```{r}
final_k_wf <- 
  k_wf %>% 
  finalize_workflow(parameters = k_best)
final_k_fit <- 
  final_k_wf %>%
  last_fit(omic_split, 
           metrics = metrics) 
final_k_fit %>%
  collect_metrics()
final_k_fit %>%
  collect_predictions() %>% 
  roc_curve(truth = STATUS, estimate = .pred_no) %>% 
  autoplot()
```

## xg boost

### model specification and workflow creation
```{r}
bt <- boost_tree(learn_rate = tune(), ) %>% 
  set_mode("classification") %>% 
  set_engine("xgboost")
bt_wf <- 
  workflow() %>% 
  add_model(bt) %>% 
  add_recipe(omic_recipe) 
```


### prepare for hyperparameter tuning
```{r}
bt_param <- parameters(bt)
bt_grid <- grid_regular(learn_rate(), levels = 3)
set.seed(525)
omic_folds <- vfold_cv(omic_train, v = 5, repeats = 1)
```


### train and tune model (with tune_grid)
```{r}
doParallel::registerDoParallel(cores = 5)
metrics = metric_set(roc_auc, accuracy)
bt_res <- 
  bt_wf %>% 
  tune_grid(
    resamples = omic_folds,
    grid = bt_grid, 
    metrics = metrics,
    control = control_grid(save_pred = TRUE)
    )
doParallel::stopImplicitCluster()
foreach::registerDoSEQ()
```


### evaluate the models
```{r}
bt_res %>% 
  collect_metrics() %>% 
  ggplot(aes(x = learn_rate, y = mean, color = .metric)) + 
  geom_point() + 
  geom_line() + ylab("") +
  ggtitle(label = "model performance", 
          subtitle = "xgboost") 
bt_best <- bt_res %>%
  select_best(metric = "roc_auc")
bt_auc <- 
  bt_res %>% 
  collect_predictions(parameters = bt_best) %>% 
  roc_curve(truth = STATUS, estimate = .pred_no) %>% 
  mutate(model = "xgboost")
autoplot(lr_auc)
```

### train and fit the best model
```{r}
final_bt_wf <- 
  bt_wf %>% 
  finalize_workflow(parameters = bt_best)
final_bt_fit <- 
  final_bt_wf %>%
  last_fit(omic_split, 
           metrics = metrics) 
final_bt_fit %>%
  collect_metrics()
final_bt_fit %>%
  collect_predictions() %>% 
  roc_curve(truth = STATUS, estimate = .pred_no) %>% 
  autoplot()
```

## svm polynomial

### model specification and workflow creation
```{r}
svmp <- svm_poly(degree = tune()) %>% 
  set_engine("kernlab") %>% 
  set_mode("classification")
svmp_wf <- 
  workflow() %>% 
  add_model(svmp) %>% 
  add_recipe(omic_recipe) 
```


### prepare for hyperparameter tuning
```{r}
svm_param <- parameters(svmp)
svmp_grid <- grid_regular(degree(), levels = 3)
set.seed(525)
omic_folds <- vfold_cv(omic_train, v = 5, repeats = 1)
```


### train and tune model (with tune_grid)
```{r}
doParallel::registerDoParallel(cores = 5)
metrics = metric_set(roc_auc, accuracy)
svmp_res <- 
  svmp_wf %>% 
  tune_grid(
    resamples = omic_folds,
    grid = svmp_grid, 
    metrics = metrics,
    control = control_grid(save_pred = TRUE)
    )
doParallel::stopImplicitCluster()
foreach::registerDoSEQ()
```


### evaluate the models
```{r}
svmp_res %>% 
  collect_metrics() %>% 
  ggplot(aes(x = degree, y = mean, color = .metric)) + 
  geom_point() + 
  geom_line() + ylab("") +
  ggtitle(label = "model performance", 
          subtitle = "svm polynomial") 
svmp_best <- svmp_res %>%
  select_best(metric = "roc_auc")
svmp_auc <- 
  svmp_res %>% 
  collect_predictions(parameters = svmp_best) %>% 
  roc_curve(truth = STATUS, estimate = .pred_no) %>% 
  mutate(model = "svm polynomial")
autoplot(svmp_auc)
```

### train and fit the best model
```{r}
final_svmp_wf <- 
  svmp_wf %>% 
  finalize_workflow(parameters = svmp_best)
final_svmp_fit <- 
  final_svmp_wf %>%
  last_fit(omic_split, 
           metrics = metrics) 
final_svmp_fit %>%
  collect_metrics()
final_svmp_fit %>%
  collect_predictions() %>% 
  roc_curve(truth = STATUS, estimate = .pred_no) %>% 
  autoplot()
```

## svm radial kernel

### model specification and workflow creation
```{r}
svmr <- svm_rbf(rbf_sigma = tune(), 
                cost = tune()) %>% 
  set_engine("kernlab") %>% 
  set_mode("classification")
svmr_wf <- 
  workflow() %>% 
  add_model(svmr) %>% 
  add_recipe(omic_recipe) 
```


### prepare for hyperparameter tuning
```{r}
svmr_param <- parameters(svmr)
svmr_grid <- 
  grid_regular(cost(), rbf_sigma(), levels = 3)
set.seed(525)
omic_folds <- vfold_cv(omic_train, v = 5, repeats = 1)
```


### train and tune model (with tune_grid)
```{r}
doParallel::registerDoParallel(cores = 5)
metrics = metric_set(roc_auc, accuracy)
svmr_res <- 
  svmr_wf %>% 
  tune_grid(
    resamples = omic_folds,
    grid = svmr_grid, 
    metrics = metrics,
    control = control_grid(save_pred = TRUE)
    )
doParallel::stopImplicitCluster()
foreach::registerDoSEQ()
```


### evaluate the models
```{r}
svmr_res %>% 
  collect_metrics() %>% 
  ggplot(aes(x = cost, y = mean, color = .metric)) + 
  geom_point() + 
  geom_line() + ylab("") +
  facet_wrap(~rbf_sigma) +
  ggtitle(label = "model performance", 
          subtitle = "svm radial")
svmr_best <- svmr_res %>%
  select_best(metric = "roc_auc")
svmr_auc <- 
  svmr_res %>% 
  collect_predictions(parameters = svmr_best) %>% 
  roc_curve(truth = STATUS, estimate = .pred_no) %>% 
  mutate(model = "svm radial")
autoplot(svmr_auc)
```

### train and fit the best model
```{r}
final_svmr_wf <- 
  svmr_wf %>% 
  finalize_workflow(parameters = svmr_best)
final_svmr_fit <- 
  final_svmr_wf %>%
  last_fit(omic_split, 
           metrics = metrics) 
final_svmr_fit %>%
  collect_metrics()
final_svmr_fit %>%
  collect_predictions() %>% 
  roc_curve(truth = STATUS, estimate = .pred_no) %>% 
  autoplot()
```

## linear discriminant

### model specification and workflow creation
```{r}
library(discrim)
ld <- discrim_linear(penalty = tune()) %>% 
  set_engine("mda") %>% 
  set_mode("classification")
ld_wf <- 
  workflow() %>% 
  add_model(ld) %>% 
  add_recipe(omic_recipe) 
```


### prepare for hyperparameter tuning
```{r}
ld_param <- parameters(ld)
ld_grid <- grid_regular(penalty(), levels = 3)
set.seed(525)
omic_folds <- vfold_cv(omic_train, v = 5, repeats = 1)
```


### train and tune model (with tune_grid)
```{r}
doParallel::registerDoParallel(cores = 5)
metrics = metric_set(roc_auc, accuracy)
ld_res <- 
  ld_wf %>% 
  tune_grid(
    resamples = omic_folds,
    grid = ld_grid, 
    metrics = metrics,
    control = control_grid(save_pred = TRUE)
    )
doParallel::stopImplicitCluster()
foreach::registerDoSEQ()
```


### evaluate the models
```{r}
ld_res %>% 
  collect_metrics() %>% 
  ggplot(aes(x = penalty, y = mean, color = .metric)) + 
  geom_point() + 
  geom_line() + ylab("") +
  ggtitle(label = "model performance", 
          subtitle = "lda") 
ld_best <- ld_res %>%
  select_best(metric = "roc_auc")
ld_auc <- 
  ld_res %>% 
  collect_predictions(parameters = ld_best) %>% 
  roc_curve(truth = omic_problems, estimate = .pred_no) %>% 
  mutate(model = "lda")
autoplot(ld_auc)
```

### train and fit the best model
```{r}
final_ld_wf <- 
  ld_wf %>% 
  finalize_workflow(parameters = ld_best)
final_ld_fit <- 
  final_ld_wf %>%
  last_fit(omic_split, 
           metrics = metrics) 
final_ld_fit %>%
  collect_metrics()
final_ld_fit %>%
  collect_predictions() %>% 
  roc_curve(truth = omic_problems, estimate = .pred_no) %>% 
  autoplot()
```
© 2021 GitHub, Inc.
Terms
Privacy
Security
Status
Docs
Contact GitHub
Pricing
API
Training
Blog
About

